Natural variations in water discharge quantity have an inconsistent impact on sediment transport capacity in subglacial channels

### 2023-10-09 Delaneyetal2023b.tex
dropped subscript c of variable w

In a transport-limited regime, sediment discharge responds to the sediment transport capacity, defined as the amount of sediment the water can carry.
Sediment transport capacity depends on the shear stress between water and the sediment it flows  over \cite{shields1936,meyer1948,engelund1967} along with the width of the channel bottom $w_c$ over which to mobilize sediment.
The shear stress  across the channel width $\tau_t$ responds to the velocity of water $v$ flowing through the channel so that
\begin{linenomath*}
  \begin{equation}
    \label{eq:tau_t}
    \tau_t \propto w_c\, v^2,
  \end{equation}
\end{linenomath*}
%
where $w_c$ is the width of the channel. Following mass conservation, the velocity of the water flowing through a  channel is
\begin{linenomath*}
  \begin{equation}
    \label{eq:v}
    v = \frac{Q}{S},
  \end{equation}
\end{linenomath*}
where $Q$ is water discharge,  and $S$ is the channel's wetted area.

@@

The response of water velocity to changing water discharge in subglacial channels differs from subaerial ones.
The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by frictional heating of water flowing through the channel \cite{rothlisberger1972}.
As a result, the subglacial channel size only evolves relatively slowly over days to months, whereas water discharge can vary over hours \cite<e.g.>{iken1986,andrews2014,nanni2020}.
Therefore, subglacial water flow behaves more like pipe flow, and changes in water discharge $Q$ are mainly accommodated with increased or decreased water velocity $v$ \cite{alley1997} because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
%It follows that an increase in water discharge $Q$ will cause the shear stress across the channel $\tau_t$ to respond with water velocity $v$, as opposed to subaerial channels where channel width $w_c$ evolves as well.

  \begin{figure}[h]
  \centering
    \includegraphics[width=0.8\linewidth]{Fig1.pdf}
    \caption{Sketch for the different responses that subglacial and subaerial channels to increased water discharge.
      Water velocity magnitudes in the subglacial, $v_g$, and subaerial, $v_f$, channels are shown by arrow length.
      $S_g$ and $S_f$ represent the wetted area in subglacial and subaerial channels, respectively.
      The subglacial channel widths $w_{cg}$ remain unchanged, while the subaerial channel widths $w_{cf}$ evolve. Shear stress $\tau$ is responsible for the mobilization of sediment.
      %Note that in the subaerial channel parameterization, channels have rectangular shapes with their width much larger than their depth (Section~\ref{sect:fluv}).
    }
    \label{fig:cartoon}
  \end{figure}
% \ian{Come back to Daniel's comments}

@@

The width-integrated shear stress is represented as
\begin{equation}
  \label{eq:tautg}
  \tau_{tg}=w_{cg}\,\tau_g,
\end{equation}

where  channel floor width $w_{cg}$ is
\begin{equation}
  \label{eq:dh2wc}
  w_{cg} = 2  \sin \frac{\beta}{2} \sqrt{\frac{2\, S_g}{\beta -\sin \beta}}.
\end{equation}

@@

To parameterize the shear stress of water flowing across sediments in the subaerial channel, the hydraulics parameterization presented in \citeA{tucker1997} is implemented.
Using mass conservation and the Darcy-Weisbach relationship as well as assuming that the channel is wide so that the hydraulic radius is well approximated by the flow depth, then
the shear stress $\tau_f$ at the river bed is
\begin{linenomath*}
  \begin{equation}
    \label{eq:DW_tau}
    \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q}{w_{cf}} \Big)^{\frac{2}{3}} \,\nabla z_c^{\frac{2}{3}},
  \end{equation}
\end{linenomath*}
where $\nabla z_c$ is the channel slope, and $f_p$ is the friction factor for subaerial channels.
Channel width $w_{cf}$ is
\begin{equation}
  \label{eq:wcf}
  w_{cf} = k \, Q^{\alpha},
\end{equation}
%
where $k$ is a constant and $\alpha=\frac{1}{2}$ is a commonly chosen exponent \cite{leopold1953}.
Water velocity, $v_f$, is given by rearranging Equation \ref{eq:tau} as
\begin{equation}
  \label{eq:vf}
  v_f = \sqrt{\frac{8\,\tau_f}{f_p\,\rho_w}}.
\end{equation}
%
As above, the width-integrated shear stress is
\begin{equation}
    \label{eq:tautf}
    \tau_{tf}=w_{cf}\,\tau_f.
  \end{equation}

@@

\appendix
\section{Table of variables, parameters and constants}
\begin{table}[h]
  \centering
  \caption{Variables, parameters, and constants used in this work.
    Where two values are given, the first refers to  \alpine{}, a case from Fieschergletscher, and the second to a glacier marginal to the \icesheet{} a case from Leverett Glacier}
  \begin{tabular}{ l  c  c c }
    Name &Symbol&  Value&Units \\ \hline
    \textbf{Variables}  & & & \\
    Water discharge  & $Q$& & $\mathrm{m^{3}\,s^{-1}}$ \\
    Water velocity (glacier, subaerial)  & $v_g,\,v_{f}$& & $\mathrm{m\,s^{-1}}$ \\
    Channel wetted area (glacier, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
    Channel depth (subaerial) & &$H$& $\mathrm{m}$\\
    Hydraulic diameter &$D_h$&&$\mathrm{m}$\\
    Width of channel floor (glacial, subaerial) & $w_{cg},w_{cf}$&  & $\mathrm{m}$     \\
    Hydraulic head &$\Delta h$&& $\mathrm{m}$\\
    Hydraulic gradient &$\Psi=\frac{\Delta h}{l}$&& $\mathrm{m\, m^{-1}}$\\

@@

\mauro{Notation: either use $h$ and $w_c$ for the river, or otherwise, somehow free $h$. Below I use $H$ and $W$ for now.}

@@

Sediment discharge is given by the MPM, EH and Bagnold formulations as
\begin{equation}
  Q_s \propto w_c\, \tau^{3/2}, \quad Q_s \propto w_c\, \tau^{5/2}, \quad Q_s \propto w_c \left(\frac{\Omega}{w_c}\right)^{3/2} H^{-2/3},
\end{equation}
respectively, for conditions well above the transport onset threshold.

The  subaerial channel is assumed to have a width $w_c$ much greater than its depth $H$, such that $D_h\approx 4h$, and to have a constant head gradient ($\Psi$) given by the topography.
Further, it is  assumed that its width can be approximated by a relation $w_c \propto Q^\alpha$ (Equation~\ref{eq:wcf})with $\alpha\in [0,1]$; typically $\alpha \approx 1/2$ (a so-called regime channel).
End members $\alpha=0$ or $\alpha=1$ correspond a  subaerial channels of constant width (a slot canyon) or depth (no natural equivalent), respectively.
%
For a steady state R-channel, it is assumed that  $\Psi$ is constant (approximated by the gradient of the Shreve \citeyear{shreve1972} potential \footnote{Note that the R-channel model presented above (Eq.~\eqref{eq:dS_dt}--\eqref{eq:dh2wc}), calculates $\Psi$ from the time evolving $S_g$ (Eq.~\eqref{eq:dS_dt}) via the Darcy-Weisbach equation~\eqref{eq:dh} and thus no Shreve approximation is then needed.}) and that $S$ adjusts such that it is in steady state with $\Psi$ and $Q$.
% Conversely, pipe flow has $S$ fixed and $\Psi$ adjusts to satisfy the given $Q$.
%
Pipe flow like conditions occur when a R-channel is subjected to rapid discharge variations such that the channel cannot adjust its size significantly on the time scale of the discharge variations.
In this case, it is assumed that the cross-sectional area $S$ is fixed, and $\Psi$ adjusts to the specified $Q$.
For both R-channel and pipe flow, it is assumed assume that $D_h \propto S^{1/2}$.

@@

Table~\ref{tab:Qs2} shows how the proxy $w_c\tau$ as well as $Q_s$ of the MPM, EH and Bagnold sediment transport formulas scale with respect to $Q$, $\Psi$ or $S$ for the three different regimes and for different $\alpha$.
Remarkably, the Bagnold formula has a negative exponent for $f$ in all but the pipe-flow regime, which seems rather unexpected.
The total transport formula EH gives a slightly stronger dependence on all variables as should be expected due to the larger exponent on $\tau$ of $\frac{5}{2}$ versus  $\frac{3}{2}$  for the others.
Albeit, the sediment transport response in pipe-flow for the Bagnold case is close to EH.
Conversely, the sediment transport proxy $w_g\,\tau_{g}, w_f\,\tau_{f}$ used within this publication scales only as $Q^2$ for pipe-flow, whereas the others scale at least with $Q^3$.
The exponent for the width scaling $\alpha$ only impacts relationship between sediment transport and water discharge in the EH relation in any meaningful way.
However, for the value of $\alpha$ around $\frac{1}{2}$ (a value appropriate for most streams), the exponent on $Q$ for EH is only slightly above what the other transport relations give.

@@

\begin{table}
  \caption{Relations for hydraulic variables for the three drainage regimes:  subaerial channels, R-channel and pipe-flow.  Darcy-Weisbach equation is abbreviated with ``D-W'', and stream power with ``Stream p.''. }
  \label{tab:eqs1}
\begin{tabular}{llllll}
  Regime & Fixed & Determined via D-W
   & Additional relations & Shear stress & Stream p.\\
& & (Equation~\ref{eq:DW})  &  &  \(\tau \propto\) & \(\Omega \propto\)\\
\hline
  Subaerial & \(\Psi\) & \(H \propto f^{1/3}\, Q^{2/3-2\alpha/3} \, \Psi^{-1/3}\) & \makecell{\(w_c\,\propto Q^\alpha\) \\ \(S=w_cH\) \\ \(D_h\propto H\)} & \(f^{1/3} Q^{2/3-2\alpha/3}  \Psi^{2/3}\) & \(Q\, \Psi\)\\
  R-channel & \(\Psi\) & \(S\, \propto f^{2/5}\, Q^{4/5} \, \Psi^{-2/5}\) & \(D_h\propto w_c \propto H \propto S^{1/2}\) & \(f^{1/5} Q^{2/5} \, \Psi^{4/5}\) & \(Q\, \Psi\)\\
Pipe & \(S\) & \(\Psi \propto f \, Q^2\, S^{-5/2}\) & \(D_h\propto w_c \propto H \propto S^{1/2}\) & \(f Q^2 S^{-2}\) & \(f\, Q^3 S^{-5/2}\)\\
\end{tabular}
\end{table}

\begin{table}
  \caption{Sediment transport proxy ($w_c\tau$) and rates for the three considered different transport formulas: MPM \cite{meyer1948}, EH \cite{engelund1967}, and Bagnold \cite{bagnold1980}.
    }
  \label{tab:Qs}
\begin{tabular}{lllll}
 & Width \(\times \,\, \tau\) & MPM & EH & Bagnold\\
 & \(w_c\, \tau\) & \(Q_s \propto w_c\, \tau^{3/2}\) & \(Q_s \propto w_c\, \tau^{5/2}\) & \(Q_s \propto w_c^{-1/2}\, \Omega^{3/2} H^{-2/3}\)\\
\hline
Subaerial  & \(f^{1/3}\, Q^{2/3+\alpha/3}\,  \Psi^{2/3}\) & \(f^{1/2}\, Q \, \Psi\) & \(f^{5/6}\, Q^{5/3 - 2\alpha/3} \, \Psi^{5/3}\) & \(f^{-2/9}\, Q^{19/18-\alpha/18} \, \Psi^{31/18}\)\\
R-channel & \(f^{2/5}\, Q^{4/5} \, \Psi^{3/5}\) & \(f^{1/2}\, Q \, \Psi\) & \(f^{7/10}\, Q^{7/5}\, \Psi^{9/5}\) & \(f^{-7/30}\, Q^{31/30}\, \Psi^{26/15}\)\\
Pipe & \(f \, Q^2 \, S^{-1}\) & \(f^{3/2}\, Q^3 \, S^{-5/2}\) & \(f^{5/2}\, Q^5\, S^{-9/2}\) & \(f^{3/2} \, Q^{9/2} \, S^{-14/3}\)\\
\end{tabular}
\end{table}

\begin{table}
  \caption{As Table~\ref{tab:Qs} but with the fractional exponents stated as rounded decimal numbers.  For the subaerial regime the exponents are for displayed for the likely width-exponent $\alpha=1$, as well as its end members 0 (slot canyon) and 1 (only width increases and not depth).
    }
  \label{tab:Qs2}
\begin{tabular}{lllll}
 & Width \(\times \,\, \tau\) & MPM & EH & Bagnold\\
 & \(w_c\, \tau\) & \(Q_s \propto w_c\, \tau^{3/2}\) & \(Q_s \propto w_c\, \tau^{5/2}\) & \(Q_s \propto w_c^{-1/2}\, \Omega^{3/2} H^{-2/3}\)\\
\hline
Subaerial (\(\alpha=0\)) & \(f^{0.7}\, Q^{0.7}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{1.7} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.1} \, \Psi^{1.7}\)\\
Subaerial (\(\alpha=0.5\)) & \(f^{0.7}\, Q^{0.8}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{1.3} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\
Subaerial (\(\alpha=1\)) & \(f^{0.7}\, Q^{1.0}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{0.7} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\[3pt]
R-channel & \(f^{0.4}\, Q^{0.8} \, \Psi^{0.6}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.7}\, Q^{1.4}\, \Psi^{1.8}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\
Pipe & \(f \,\quad Q^{2\phantom{.0}} \, S^{-1}\) & \(f^{1.5}\, Q^3 \, S^{-2.5}\) & \(f^{2.5}\, Q^{5\phantom{.0}}\, S^{-4.5}\) & \(f^{1.5} \,\,\,\,\, Q^{4.5} \, S^{-4.7}\)\\
\end{tabular}
\end{table}

>>>>

In a transport-limited regime, sediment discharge responds to the sediment transport capacity, defined as the amount of sediment the water can carry.
Sediment transport capacity depends on the shear stress between water and the sediment it flows  over \cite{shields1936,meyer1948,engelund1967} along with the width of the channel bottom $w$ over which to mobilize sediment.
The shear stress  across the channel width $\tau_t$ responds to the velocity of water $v$ flowing through the channel so that
\begin{linenomath*}
  \begin{equation}
    \label{eq:tau_t}
    \tau_t \propto w\, v^2,
  \end{equation}
\end{linenomath*}
%
where $w$ is the width of the channel. Following mass conservation, the velocity of the water flowing through a  channel is
\begin{linenomath*}
  \begin{equation}
    \label{eq:v}
    v = \frac{Q}{S},
  \end{equation}
\end{linenomath*}
where $Q$ is water discharge,  and $S$ is the channel's wetted area.

@@

The response of water velocity to changing water discharge in subglacial channels differs from subaerial ones.
The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by frictional heating of water flowing through the channel \cite{rothlisberger1972}.
As a result, the subglacial channel size only evolves relatively slowly over days to months, whereas water discharge can vary over hours \cite<e.g.>{iken1986,andrews2014,nanni2020}.
Therefore, subglacial water flow behaves more like pipe flow, and changes in water discharge $Q$ are mainly accommodated with increased or decreased water velocity $v$ \cite{alley1997} because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
%It follows that an increase in water discharge $Q$ will cause the shear stress across the channel $\tau_t$ to respond with water velocity $v$, as opposed to subaerial channels where channel width $w$ evolves as well.

  \begin{figure}[h]
  \centering
    \includegraphics[width=0.8\linewidth]{Fig1.pdf}
    \caption{Sketch for the different responses that subglacial and subaerial channels to increased water discharge.
      Water velocity magnitudes in the subglacial, $v_g$, and subaerial, $v_f$, channels are shown by arrow length.
      $S_g$ and $S_f$ represent the wetted area in subglacial and subaerial channels, respectively.
      The subglacial channel widths $w_g$ remain unchanged, while the subaerial channel widths $w_f$ evolve. Shear stress $\tau$ is responsible for the mobilization of sediment.
      %Note that in the subaerial channel parameterization, channels have rectangular shapes with their width much larger than their depth (Section~\ref{sect:fluv}).
    }
    \label{fig:cartoon}
  \end{figure}
% \ian{Come back to Daniel's comments}

@@

The width-integrated shear stress is represented as
\begin{equation}
  \label{eq:tautg}
  \tau_{tg}=w_g\,\tau_g,
\end{equation}

where  channel floor width $w_g$ is
\begin{equation}
  \label{eq:dh2wc}
  w_g = 2  \sin \frac{\beta}{2} \sqrt{\frac{2\, S_g}{\beta -\sin \beta}}.
\end{equation}

@@

To parameterize the shear stress of water flowing across sediments in the subaerial channel, the hydraulics parameterization presented in \citeA{tucker1997} is implemented.
Using mass conservation and the Darcy-Weisbach relationship as well as assuming that the channel is wide so that the hydraulic radius is well approximated by the flow depth, then
the shear stress $\tau_f$ at the river bed is
\begin{linenomath*}
  \begin{equation}
    \label{eq:DW_tau}
    \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q}{w_f} \Big)^{\frac{2}{3}} \,\nabla z_c^{\frac{2}{3}},
  \end{equation}
\end{linenomath*}
where $\nabla z_c$ is the channel slope, and $f_p$ is the friction factor for subaerial channels.
Channel width $w_f$ is
\begin{equation}
  \label{eq:wcf}
  w_f = k \, Q^{\alpha},
\end{equation}
%
where $k$ is a constant and $\alpha=\frac{1}{2}$ is a commonly chosen exponent \cite{leopold1953}.
Water velocity, $v_f$, is given by rearranging Equation \ref{eq:tau} as
\begin{equation}
  \label{eq:vf}
  v_f = \sqrt{\frac{8\,\tau_f}{f_p\,\rho_w}}.
\end{equation}
%
As above, the width-integrated shear stress is
\begin{equation}
    \label{eq:tautf}
    \tau_{tf}=w_f\,\tau_f.
  \end{equation}

@@

\appendix
\section{Table of variables, parameters and constants}
\begin{table}[h]
  \centering
  \caption{Variables, parameters, and constants used in this work.
    Where two values are given, the first refers to  \alpine{}, a case from Fieschergletscher, and the second to a glacier marginal to the \icesheet{} a case from Leverett Glacier}
  \begin{tabular}{ l  c  c c }
    Name &Symbol&  Value&Units \\ \hline
    \textbf{Variables}  & & & \\
    Water discharge  & $Q$& & $\mathrm{m^{3}\,s^{-1}}$ \\
    Water velocity (glacier, subaerial)  & $v_g,\,v_{f}$& & $\mathrm{m\,s^{-1}}$ \\
    Channel wetted area (glacier, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
    Channel depth (subaerial) & &$H$& $\mathrm{m}$\\
    Hydraulic diameter &$D_h$&&$\mathrm{m}$\\
    Width of channel floor (glacial, subaerial) & $w_g,w_f$&  & $\mathrm{m}$     \\
    Hydraulic head &$\Delta h$&& $\mathrm{m}$\\
    Hydraulic gradient &$\Psi=\frac{\Delta h}{l}$&& $\mathrm{m\, m^{-1}}$\\

@@

\mauro{Notation: either use $h$ and $w$ for the river, or otherwise, somehow free $h$. Below I use $H$ and $W$ for now.}

@@

Sediment discharge is given by the MPM, EH and Bagnold formulations as
\begin{equation}
  Q_s \propto w\, \tau^{3/2}, \quad Q_s \propto w\, \tau^{5/2}, \quad Q_s \propto w \left(\frac{\Omega}{w}\right)^{3/2} H^{-2/3},
\end{equation}
respectively, for conditions well above the transport onset threshold.

The  subaerial channel is assumed to have a width $w$ much greater than its depth $H$, such that $D_h\approx 4h$, and to have a constant head gradient ($\Psi$) given by the topography.
Further, it is  assumed that its width can be approximated by a relation $w \propto Q^\alpha$ (Equation~\ref{eq:wcf})with $\alpha\in [0,1]$; typically $\alpha \approx 1/2$ (a so-called regime channel).
End members $\alpha=0$ or $\alpha=1$ correspond a  subaerial channels of constant width (a slot canyon) or depth (no natural equivalent), respectively.
%
For a steady state R-channel, it is assumed that  $\Psi$ is constant (approximated by the gradient of the Shreve \citeyear{shreve1972} potential \footnote{Note that the R-channel model presented above (Eq.~\eqref{eq:dS_dt}--\eqref{eq:dh2wc}), calculates $\Psi$ from the time evolving $S_g$ (Eq.~\eqref{eq:dS_dt}) via the Darcy-Weisbach equation~\eqref{eq:dh} and thus no Shreve approximation is then needed.}) and that $S$ adjusts such that it is in steady state with $\Psi$ and $Q$.
% Conversely, pipe flow has $S$ fixed and $\Psi$ adjusts to satisfy the given $Q$.
%
Pipe flow like conditions occur when a R-channel is subjected to rapid discharge variations such that the channel cannot adjust its size significantly on the time scale of the discharge variations.
In this case, it is assumed that the cross-sectional area $S$ is fixed, and $\Psi$ adjusts to the specified $Q$.
For both R-channel and pipe flow, it is assumed assume that $D_h \propto S^{1/2}$.

@@

Table~\ref{tab:Qs2} shows how the proxy $w\tau$ as well as $Q_s$ of the MPM, EH and Bagnold sediment transport formulas scale with respect to $Q$, $\Psi$ or $S$ for the three different regimes and for different $\alpha$.
Remarkably, the Bagnold formula has a negative exponent for $f$ in all but the pipe-flow regime, which seems rather unexpected.
The total transport formula EH gives a slightly stronger dependence on all variables as should be expected due to the larger exponent on $\tau$ of $\frac{5}{2}$ versus  $\frac{3}{2}$  for the others.
Albeit, the sediment transport response in pipe-flow for the Bagnold case is close to EH.
Conversely, the sediment transport proxy $w_g\,\tau_{g}, w_f\,\tau_{f}$ used within this publication scales only as $Q^2$ for pipe-flow, whereas the others scale at least with $Q^3$.
The exponent for the width scaling $\alpha$ only impacts relationship between sediment transport and water discharge in the EH relation in any meaningful way.
However, for the value of $\alpha$ around $\frac{1}{2}$ (a value appropriate for most streams), the exponent on $Q$ for EH is only slightly above what the other transport relations give.

@@

\begin{table}
  \caption{Relations for hydraulic variables for the three drainage regimes:  subaerial channels, R-channel and pipe-flow.  Darcy-Weisbach equation is abbreviated with ``D-W'', and stream power with ``Stream p.''. }
  \label{tab:eqs1}
\begin{tabular}{llllll}
  Regime & Fixed & Determined via D-W
   & Additional relations & Shear stress & Stream p.\\
& & (Equation~\ref{eq:DW})  &  &  \(\tau \propto\) & \(\Omega \propto\)\\
\hline
  Subaerial & \(\Psi\) & \(H \propto f^{1/3}\, Q^{2/3-2\alpha/3} \, \Psi^{-1/3}\) & \makecell{\(w\,\propto Q^\alpha\) \\ \(S=wH\) \\ \(D_h\propto H\)} & \(f^{1/3} Q^{2/3-2\alpha/3}  \Psi^{2/3}\) & \(Q\, \Psi\)\\
  R-channel & \(\Psi\) & \(S\, \propto f^{2/5}\, Q^{4/5} \, \Psi^{-2/5}\) & \(D_h\propto w \propto H \propto S^{1/2}\) & \(f^{1/5} Q^{2/5} \, \Psi^{4/5}\) & \(Q\, \Psi\)\\
Pipe & \(S\) & \(\Psi \propto f \, Q^2\, S^{-5/2}\) & \(D_h\propto w \propto H \propto S^{1/2}\) & \(f Q^2 S^{-2}\) & \(f\, Q^3 S^{-5/2}\)\\
\end{tabular}
\end{table}

\begin{table}
  \caption{Sediment transport proxy ($w\tau$) and rates for the three considered different transport formulas: MPM \cite{meyer1948}, EH \cite{engelund1967}, and Bagnold \cite{bagnold1980}.
    }
  \label{tab:Qs}
\begin{tabular}{lllll}
 & Width \(\times \,\, \tau\) & MPM & EH & Bagnold\\
 & \(w\, \tau\) & \(Q_s \propto w\, \tau^{3/2}\) & \(Q_s \propto w\, \tau^{5/2}\) & \(Q_s \propto w^{-1/2}\, \Omega^{3/2} H^{-2/3}\)\\
\hline
Subaerial  & \(f^{1/3}\, Q^{2/3+\alpha/3}\,  \Psi^{2/3}\) & \(f^{1/2}\, Q \, \Psi\) & \(f^{5/6}\, Q^{5/3 - 2\alpha/3} \, \Psi^{5/3}\) & \(f^{-2/9}\, Q^{19/18-\alpha/18} \, \Psi^{31/18}\)\\
R-channel & \(f^{2/5}\, Q^{4/5} \, \Psi^{3/5}\) & \(f^{1/2}\, Q \, \Psi\) & \(f^{7/10}\, Q^{7/5}\, \Psi^{9/5}\) & \(f^{-7/30}\, Q^{31/30}\, \Psi^{26/15}\)\\
Pipe & \(f \, Q^2 \, S^{-1}\) & \(f^{3/2}\, Q^3 \, S^{-5/2}\) & \(f^{5/2}\, Q^5\, S^{-9/2}\) & \(f^{3/2} \, Q^{9/2} \, S^{-14/3}\)\\
\end{tabular}
\end{table}

\begin{table}
  \caption{As Table~\ref{tab:Qs} but with the fractional exponents stated as rounded decimal numbers.  For the subaerial regime the exponents are for displayed for the likely width-exponent $\alpha=1$, as well as its end members 0 (slot canyon) and 1 (only width increases and not depth).
    }
  \label{tab:Qs2}
\begin{tabular}{lllll}
 & Width \(\times \,\, \tau\) & MPM & EH & Bagnold\\
 & \(w\, \tau\) & \(Q_s \propto w\, \tau^{3/2}\) & \(Q_s \propto w\, \tau^{5/2}\) & \(Q_s \propto w^{-1/2}\, \Omega^{3/2} H^{-2/3}\)\\
\hline
Subaerial (\(\alpha=0\)) & \(f^{0.7}\, Q^{0.7}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{1.7} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.1} \, \Psi^{1.7}\)\\
Subaerial (\(\alpha=0.5\)) & \(f^{0.7}\, Q^{0.8}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{1.3} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\
Subaerial (\(\alpha=1\)) & \(f^{0.7}\, Q^{1.0}\,  \Psi^{0.7}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.8}\, Q^{0.7} \, \Psi^{1.7}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\[3pt]
R-channel & \(f^{0.4}\, Q^{0.8} \, \Psi^{0.6}\) & \(f^{0.5}\, Q \,\,\,\, \Psi\) & \(f^{0.7}\, Q^{1.4}\, \Psi^{1.8}\) & \(f^{-0.2}\, Q^{1.0} \, \Psi^{1.7}\)\\
Pipe & \(f \,\quad Q^{2\phantom{.0}} \, S^{-1}\) & \(f^{1.5}\, Q^3 \, S^{-2.5}\) & \(f^{2.5}\, Q^{5\phantom{.0}}\, S^{-4.5}\) & \(f^{1.5} \,\,\,\,\, Q^{4.5} \, S^{-4.7}\)\\
\end{tabular}
\end{table}
